% gives a short introduction on what the IDL file is
% what the generated code looks like and how to use
% it
\chapter{Interface Definitions}
\label{section1}

The use of IDL compilers is well established 
in distributed systems. An IDL compiler provides you
a level of abstraction for the communication with a 
remote service.

Assuming you want to write a server for L4 manually you
have to solve several problems. You have to communicate
your request from a client to the server, which means that you
build a message and send an IPC containing this message
to the server. The server has to wait for request,
extract the data from the message and dispatch the
request to the appropriate function, which does the
requested work. After this function returns, the server
has to pack the response into another message and reply
to the client.

With the help of an IDL compiler we can automate all of the
above steps --- except the implementation of the server
function. This can be done using a description of the 
server's interface which is usually written using an
interface description language (IDL).

Currently, \dice{} supports two different IDLs --- CORBA 
IDL and DCE IDL. Since we think that DCE IDL is more
powerful, we added options to express
L4 specific properties to the DCE IDL. These attributes
are described in detail in section~\ref{sec:attributes}.

For the CORBA IDL there exist multiple well-defined 
language mappings. Since these are widely accepted,
\dice{} generates code that conforms to the CORBA C
language mapping \cite{corba-clm}.

\section{A Simple Interface}

To illustrate the use of \dice{} we show a simple example
here. The language of our IDL examples is DCE IDL. See
the appendix for a CORBA IDL version of the example.

\subsection{Interface Description Language}
An interface description file consists of at least an {\em interface
specification} which includes one or more {\em function declarations}.
As described a client may send data to the server and
receive a response containing data. We have to differentiate
these different kinds of data from each other. Data sent to
the server has the attribute \verb|in| and data sent back to
the client has the attribute \verb|out|.

\begin{verbatim}
interface simple
{
  void foo([in] int parameter);
}
\end{verbatim}

The above example sends a simple parameter to the server and
does not require any data in the the answer. Nonetheless does
the client continue its work only after it received an (empty) 
answer from the server. These calls are always synchronous.
Asynchronous calls are described in detail in section~\ref{sec:asynchronous}.

\section{Supported Types}

\subsection{Integer Types}
The DCE IDL supports the following types:

\begin{center}
\begin{longtable}{|l|l|p{6cm}|}%
  \multicolumn{3}{r}{next page \dots}\\\endfoot%
  \multicolumn{3}{l}{\dots continued from last page}\\\endhead%
  \endfirsthead%
  \endlastfoot%
  \hline
type & size & value range \\ \hline
$[$ signed $]$ small $[$ int $]$ & 8 bit & -128... 128 \\
unsigned small $[$ int $]$ & 8 bit & 0... 255 \\ 
$[$ signed $]$ short $[$ int $]$ & 16 bit & -32'768... 32'767 \\
unsigned short $[$ int $]$ & 16 bit & 0... 65'565 \\
signed & 32 bit & -2'147'483'648... 2'147'483'647 \\
$[$ signed $]$ int & 32 bit & -2'147'483'648... 2'147'483'647 \\
$[$ signed $]$ long $[$ int $]$ & 32 bit & -2'147'483'648... 2'147'483'648 \\
unsigned $[$ int $]$ & 32 bit & 0... 4'294'967'295 \\
unsigned long $[$ int $]$ & 32 bit & 0... 4'294'967'295 \\
$[$ signed $]$ long long $[$ int $]$ & 64 bit & -9'223'372'036'854'775'808... 9'223'372'036'854'775'807 \\
unsigned long long $[$ int $]$ & 64 bit & 0... 18'446'744'073'709'551'615 \\
$[$ signed $]$ hyper $[$ int $]$ & 64 bit & -9'223'372'036'854'775'808... 8'223'372'036'854'775'807 \\
unsigned hyper $[$ int $]$ & 64 bit & 0... 18'446'744'073'709'551'615 \\
\hline
\end{longtable}
\end{center}

The CORBA IDL supports the following types:

\begin{center}
\begin{tabular}{|l|l|p{5cm}|}
\hline
type & size & value range \\ \hline
short & 16 bit & -32'768... 32'767 \\
unsigned short & 16 bit & 0... 65'565 \\
long & 32 bit & -2'147'483'648... 2'147'483'648 \\
unsigned long & 32 bit & 0... 4'294'967'295 \\
long long & 64 bit & -9'223'372'036'854'775'808... 9'223'372'036'854'775'807 \\
unsigned long long & 64 bit & 0... 18'446'744'073'709'551'615 \\
\hline
\end{tabular}
\end{center}

\subsection{Floating Point Types}
The DCE and CORBA IDL support the following types:

% TODO: precision

\begin{center}
\begin{tabular}{|l|l|}
\hline
type & size \\ \hline
float & 32 bit \\
double & 64 bit \\
long double & 80 bit \\
\hline
\end{tabular}
\end{center}

\subsection{Other Types}

\dice{} also supports miscellaneous types for DCE IDL:

\begin{center}
\begin{tabular}{|l|l|l|}
\hline
type & size & values \\
\hline
byte & 8 bit & 0... 255 \\
void & undefined & none \\
unsigned char & 8 bit & 0... 255 \\
$[$ signed $]$ char & 8 bit & -128... 127 \\
boolean & 8 bit & {\tt true}, {\tt false} \\
\hline
\end{tabular}
\end{center}

The type {\tt boolean} is mapped to a {\tt unsigned char} type and
therefore has a size of 8 bit. Its values are {\tt false} if zero
and {\tt true} otherwise.

\dice{} supports the following miscellaneous types for the
CORBA IDL:

\begin{center}
\begin{tabular}{|l|l|l|}
\hline
type & size & values \\
\hline
char & 8 bit & -128... 127 \\
wchar & 16 bit & -32'768... 32'767 \\
boolean & 8 bit & {\tt true}, {\tt false} \\
octet & 8 bit & 0... 255 \\
\hline
\end{tabular}
\end{center}

The types {\tt char} and {\tt wchar} have a special semantic in the
CORBA IDL. Even though they might contains integer values, they are
interpreted as characters. The CORBA specification allows the communication
code to tranform character sets and thereby change the integer value of
a character. You should use the {\tt octet} value instead.

\subsection{L4 specific types}
There are some L4 specific types, which have been added to the
IDLs to support L4 specific semantics.
One of these types is the flexpage type, which expresses the
mapping of memory pages from one address space into another.
A flexpage can be transmitted using the {\tt flexpage} (or {\tt fpage})
type.

\begin{verbatim}
void map([in] fpage page);
\end{verbatim}

\section{Constructed Types}

\dice{} transmits constructed types as well. You may either include
or import a type via the mechanisms described in section~\ref{sec:import}
or you define it within the IDL file. The latter approach has the
advantage that you can give hints to the IDL compiler on how to
transmit the type. The disadvantage is, that you have to either
use the generated data type in your wrapper code or cast your own 
data type to the generated data type when using the generated stubs.

\subsection{Aliased Types}
You may iuse {\tt typedef} to define alias names for types, which
can be used as types of parameters or members of other constructed 
types.

\begin{verbatim}
typedef int buffer50[50];
\end{verbatim}

The specified types are provided in the generated header files and
can therefore be used in your code.

\subsection{Arrays and Sequences}
You may specify arrays in your IDL files.

Arrays are denoted by brackets, and may contain lower and upper bounds.
You may also specify a veriable sized array by using a pointer instead
of an empty pair of brackets.

The following example descibes an array with 20 elements of type \verb|long|.
There are different ways to express this. As mentioned you may specify 
lower bounds for arrays. The generated stub code will then check transmit
the data starting with the lower bound of the array ({\em currently
not implemented}). 

\begin{verbatim}
void array1([in] long param[20]);
void array2([in] long param[0..20]);
void array3([in] long param[10..30]);
\end{verbatim}

The following example shows some possibilities to specify variable
sized arrays. Variable sized arrays may have a fixed size, but the 
number of transmitted elements is determined at run-time. ({\em The
shown examples will not work, because the \verb|size_is| attribute
is missing.})

\begin{verbatim}
void var_array1([in] long param[*]);
void var_array2([in] long param[]);
void var_array3([in] long *param);
\end{verbatim}

In a CORBA IDL file you may use the {\tt sequence} keyword to define
array types. The CORBA C Language Mapping defines that {\tt sequence}
is translated into a {\tt struct}. Consider the following example:
\begin{verbatim}
IDL:
typedef sequence<long, 10> my_long_10;

C:
typedef struct
{
  unsigned long _maximum;
  unsigned long _length;
  long *_buffer;
} my_long_10;
\end{verbatim}

You may have noticed that the boundary of the sequence is silently dropped
when converting the sequence to the C type. You are responsible for enforcing
the boundary yourself\footnote{If you think this is unfair, please contact the
OMG.}.

The syntax of CORBA allows you to specify a sequence as type of a parameter.
When mapping this to C, this will produce a struct definition inside a
function declaration. This implies that the struct type is only known inside
the funtction, which means you cannot declare a variable of this type outside
the function.

\subsubsection{Strings}
Strings are regarded as a variable sized array of characters, which is 
zero-terminated.
This allows to omit the specification of length attributes for the 
parameter. Instead you specify an \verb|string| attribute. The generated
code will use string functions to determine the length of the string.

\begin{verbatim}
void hello([in, string] char* world);
\end{verbatim}

In CORBA IDL you may specify strings with boundary (similar to the sequence
type). The CORBA C Language Mapping defines to map strings and wide-strings
to \verb|char*| and \verb|wchar*| respectively, ignoring the boundaries. This
implies that you may have to enforce thes boundaries yourself.

\begin{verbatim}
IDL:
typedef string<25> my_name;

C:
typedef char* my_name;
\end{verbatim}

\subsection{Structured Types}
Within the IDL file you may specify a structured type --- a \verb|struct|
 --- the same way as you would within a C/C++ header file. Furthermore
it is possible to specify attributes with each member of the defined
\verb|struct|. One of the allowed attributes is \verb|size_is|, which
is described in more detail in Section~\ref{sec:attributes}. It allows
you to specify the run-time size of an array. The definition of such a
structure could look like this:
\begin{verbatim}
struct _string
{
  [size_is(length), string] char* buffer;
  unsigned length;
};
\end{verbatim}
This is similar to the usage as parameters:
\begin{verbatim}
[in, size_is(length), string] char* buffer,
[in] unsigned length
\end{verbatim}

% Valid attriutes are:
% ...

A \verb|struct| is mostly marshaled by copying it as is into the 
message buffer. If the \verb|struct| contains variable sized members,
such as the \verb|buffer| member of the above example, these are
marshaled separately after the parameter of the structured type.

\subsubsection{Bitfields}
You may specify bitfields in a structured type. Consider that this 
is highly platform dependent and the usage of bitfields may lead
to unwanted results.

\begin{verbatim}
struct {
  unsigned int first : 3;
  unsigned int second : 5;
  unsigned int third : 8;
};
\end{verbatim}

\subsection{Unions}
Within your IDL specification you
may define a union in C syntax or in IDL syntax. The difference is made
when marshaling the union into the message buffer. A ``C-style'' union
is simply copied into the message buffer. Therefore it will consume the
space of the largest of its members.

The ``IDL-style'' union declares a decision making variable, which is then
used to select the member to transmit. Such a union could look like this:
\begin{verbatim}
union switch(long which_member) _union
{
case 1:
  long l_mem;
case 2:
  double d_mem;
case 3:
case 4:
  long array[100];
default:
  byte status;
};
\end{verbatim}
If the variable \verb|which_member| has the value 1, the member \verb|l_mem|
is transmitted. When defining a switch variable and the case statement, you
have to consider that they are used in a C switch statement for comparison.
Thus it is valid to use a \verb|char s_var| and define the cases with character
values.

The generated code for an ``IDL-style'' union includes a switch statement to 
decide which member to transmit. This may be relevant if the union may
tremendously vary in size and a comparison would possibly minimize the
amount of data to be copied. But, such a union will always need the switch
variable to be transmitted as well --- at the receiver's side there is 
also a switch statement which decides what to unmarshal from the message 
buffer.

A union may also have constructed or variable sized members. Members of
variable size are marshalled after the union.

\section{Deriving Interfaces}
You can derive an interface from another interface. This is done
using similar syntax as with C++ classes. You name the interface to derive
followed by a colon and then the list of base interfaces.
\begin{verbatim}
interface derived : base1, base2
{
...
\end{verbatim}

The main difference to C++ class derivation is, that you mostly use derived 
classes to overload functionality of the base classes. Since there is no
implementation in an IDL file, which could be overloaded, the main purpose
of interface derivation is the expansion of the base interface's functionality.
This is done by specifying new functions in the derived interface.

You may use interface derivation to unify multiple interface definitions
into one interface. The derived interface may be empty. 
\begin{verbatim}
interface all_in_one : 
  base1, base2, 
  another_scope::base3
{};
\end{verbatim}

The generated server loop will be able to receive message from \verb|base1|,
\verb|base2|, and \verb|another_scope::base3|. The server loop
distinguishes the different function calls from another as described
in the following section (Section~\ref{sec:opcode}).

\section{Constants}
You may specify constant by either using the \verb|#define| statement
or a {\tt const} declaration. The \verb|#define| statement is parsed
by the pre-processor and will not appear in the generated code. The
{\tt const} declaration will appear in the generated code as they have
been defined in the IDL file.

\begin{verbatim}
const int foo = 4;
\end{verbatim}

You may use the declared {\tt const} variables when specifying the sizes
of arrays, etc.

\section{Attributes}
\label{sec:attributes}

This section describes some attributes, which are available
with the DCE IDL. If an attribute is available for CORBA IDL as well
this is specifically mentioned.

Attributes are used to propagate knowledge about the target
environment or usage context of the IDL specification to \dice{}.
Some attributes are used to generate code optimized for a
specific platform or architecture. Others will influence the
generation of the target code to exploit features of a specific 
L4 ABI. Therefore the attributes are important to provide
\dice{} with knowledge about the IDL specification and optimization
potential.

However, there are attributes which might be ignored on some platforms.

\subsection{Array Attributes}
Attributes describing parameters of an array mostly end on
\verb|_is| and usually take a parameter. They are:
\verb|first_is|, \verb|last_is|, \verb|length_is|, \verb|min_is|,
\verb|max_is|, and \verb|size_is|. Currently only \verb|length_is|,
\verb|size_is|, and \verb|max_is| are supported.

Starting with the latter, the attribute is used to determine
the maximum size of an array. For example do
\begin{verbatim}
[in] int parameter[100]
\end{verbatim}
and
\begin{verbatim}
[in, max_is(100)] int parameter[]
\end{verbatim}
express the same semantics.

You may also specify a parameter or constant as the 
argument of the attribute.

The \verb|length_is| or \verb|size_is| attributes are used
to determine the length of a variable sized array at run-time.
The generated code will transmit only the specified number of
elements instead of the maximum size. The \verb|str| parameter
does not have a maximum size to keep the example simple. A maximum
size will enable the server loop or client stubs to know the
size of memory to allocate for the array.

The example:
\begin{verbatim}
[in, size_is(len)] char str[], 
[in] int len
\end{verbatim}
specifies a variable sized array of bytes, which has a length
of \verb|len|. \verb|len| has to be either a parameter or
a constant. \dice{} first searches for parameters with the
secified name and, if it does not find one, it searches for
a constant with that name. An error is generated if neither 
is given.

%TODO: exact semantics of _is attributes
%TODO: are there other attributes relevant for arrays

\subsection{Strings}
As described above is it possible to define the run-time length
of a variable sized array using the \verb|length_is| or 
\verb|size_is| attribute. This may also be used for strings.
If you know that the parameter you will transmit is a zero-terminated string,
you can also let the generated code calculate the size of
the string. You simply specify the \verb|string| attribute.

\begin{verbatim}
[in, string] char* str
\end{verbatim}

In this example the \verb|str| parameter does not have a maximum
size, because it can be any size. 

\subsection{Indirect Parts IPC}
L4 allows you to transmit any data using 
indirect part IPC. You specify the address and size 
of the original data and at receiver's side the address and 
size of the receive buffer and the kernel copies the data
directly. Without indirect parts the data is copied from the
source address to the message buffer, the kernel copies the message buffer
from sender's side to the receiver's message buffer,
and the receiver copies the date from the message buffer to the target variables.
The downside of the indirect part IPC is, that the
kernel may have to establish additional temporary mappings 
of the data areas.

To make \dice{} use indirect part IPC for a specific parameter
you may use the \verb|ref| attribute. You can either combine it
with the \verb|size_is| attribute or the \verb|string| attribute
to specify the size of a variable sized array.

\begin{verbatim}
[in, size_is(size), ref] long data[],
[in] long size
\end{verbatim}

You may also transmit any other kind of data using indirect 
part IPC. Only specify the \verb|ref| attribute with the respective
parameter and the size of the data to be transmitted. 

\subsection{Receiving Indirect Parts}
The CORBA C Language Mapping specifies for the reception of variable sized
parameters the allocation of memory using the \verb|CORBA_alloc| function.
If you know the receive buffer for the variable sized parameter before
calling the client stub, you may want to give this buffer to the client
stub, so the data is copied directly into this buffer. This can be done
using \verb|prealloc| attribute:
\begin{verbatim}
[out, prealloc, size_is(len)] char** str,
[out] int len
\end{verbatim}

{\em The Section~\ref{sec:mem-indirect-parts} contains further information
on the memory management of indirect parts.}

\subsection{Error Function}
Usually the server ignores IPC errors when sending a reply to a client or
waiting for a new request. Sometimes it is useful to use timeouts when
waiting for new request to ``do something else'' meanwhile. The wait timeout 
can be specified in the \verb|CORBA_Environment| parameter of the
server loop. But how do you know when the timeout has been triggered and
what does the server loop do?

For this scenario, there exists an attribute --- \verb|error_function| --- which
makes the server loop call the specified function on every IPC error.
This allows you to implement a specific behavior for the IPC error. The
called function takes only one parameter --- an \verb|l4_msgdope_t| variable.
This can be used to check the origin of the error and take appropriate
measures.

% XXX Todo
% The other parameter is the message buffer, which should be transmitted and
% the current \verb|CORBA_Environment|. These can be used to obtain some
% context information of the error

% describes some L4 specific scenarios and how to make
% the IDL compiler do what you want

\subsection{Using other types for transmition}
Sometimes it is convenient to use other types to transmit the data,
than are used when calling a function. An example is the usage of
a constructed type, which contains variable sized members, which
are known only at run-time. If this type is declared in a C header
file, it is not possible to add attributes to the variable sized
member to define the run-time size. 

The attribute \verb|transmit_as| lets you define the type to use
when transmitting the data:
\begin{verbatim}
typedef [transmit_as(unsigned long)] 
my_C_union idl_union;
\end{verbatim}
This example will use the type \verb|unsigned long| instead of
\verb|my_C_union| when transmitting parameters of type \verb|idl_union|.

The example:
\begin{verbatim}
[in, transmit_as(unsigned char),
 size_is(size)] var_struct *t,
[in] unsigned size
\end{verbatim}
will transmit \verb|t| as a variable sized array of type \verb|unsigned char|
with size \verb|size| instead of \verb|var_struct|.

{\em Consider that the latter example requires the specification of
a pointered parameter. Without the \verb|transmit_as| attribute this
woud have the semantic of an array of type \verb|var_struct| and
\verb|size| members. But with the \verb|transmit_as| attribute the
parameter \verb|t| is considered a variable sized array with 
\verb|size| elements of type \verb|unsigned char|.}

\section{Imported and Included Files}
\label{sec:import}

\dice{} provides you with two mechanisms to include other IDL or
C/C++ files into your IDL file. The first is the well known 
preprocessor directive \verb|#include|. The second is the \verb|import|
statement. Both are resolved by \dice{}'s preprocessor.
Including a file using \verb|import| iimplies that the content 
of this file is only scanned for IDL syntax. This includes
C/C++ style type and constant defintions, but excludes C style 
function declarations and definitions.

\dice{} automatically scans the file \verb|dice-corba-types.h|
to know all the CORBA types. This file is searched in L4 tree's
include path and the include path where dice header's might be
installed to. To avoid inclusion of unwanted header files the
two include paths contain the \verb|/dice| subdir.
If you receive the error '\verb|dice-corba-types.h| cannot
be found', please check if your \verb|$(L4DIR)/include/dice| or
\verb|/usr/local/include/dice| contain that file. If not, fix
your installation or run \verb|make| in the \dice{} source
directory.

\subsection{C/C++ Attributes}
If you specify attributes (\verb|__attribute__((...))|)
with a type definition, they are ignored by the IDL compiler. This is
important if you specify attributes responsible for the memory
layout of a data type. Thus, if you define C types using attributes
check the generated C code to see if you data type is really copied
the way you intended.

\section{Message Passing}
\label{sec:message-passing}
Traditional RPC semantics include the sequence of actions of sending a
request to a server, processing the request and waiting for a reply.
Opposed to that exist programming languages, which use communication
mechanisms to synchronize or exchange data at specific points in the
program. In Ada these points are called ``rendevouz''. With this technique
it is possible to write programs, where activity is associated the
flow of data.

To provide the level of abstraction we introduced for RPCs for message
passing also, we extended the DCE IDL to let \dice{} generate simple
communication stubs.

A ``rendevouz'' point is given on L4 when one thread waits for a
specific message and another thread is sending this message to the
waiting thread. Since a message is defined in the IDL by an operation of an 
interface, we associate an attribute with the operation to denote 
the message passing semantic. It is possible for a message to be
consumed by a thread or emitted by it. This is similar to the semantic
for parameters of a RPC. Thus we used the \verb|in| and \verb|out|
attribute for message passing.

\begin{verbatim}
interface mp
{
  [in] void send(int param);
  [out] int receive(int *param);
};
\end{verbatim}

The parameters of an \verb|in| operation are automatically \verb|in| as
well. The usage of return types other than void or \verb|out| parameters
will cause an error. Accordingly are all parameters of an \verb|out|
operation \verb|out| themselves.

\dice{} generates a different set of functions for message passing
operations. At the sender's side\footnote{Beware, this can be the client's
side and the server's side.} the functions and on \verb|_send|, because
they only involve the send phase of the IPC. At the receiver's side
there are two functions, one ending on \verb|_wait| and the other ending
on \verb|_recv|. These functions receive the specified message from
any sender or a specific sender respectively. For the receiver's side
there also exists an \verb|_unmarshal| function.

{\em Currently the receiving functions are buggy.}

If the receiver's side is also the server's side the message
can also be received by a server loop. For this case \dice{} also
generates the beforementioned functions (\verb|_unmarshal|, 
\verb|_component|). 

A synonym for a function's \verb|in| attribute is the \verb|oneway|
attribute, known from DCE IDL.

\section{Asynchronous Calls}
\label{sec:asynchronous}

\section{Thoughts About \dice{} Specific Types}
When using IDL types, they mostly can be mapped to L4 specific 
message types using attributes. An indirect part is for 
instance denoted by the DCE \verb|ref| attribute. It thereby
follows the semantic declared in the DCE IDL specification.

However, there are some L4 specific IPC semantics, which
are hard to express using attributes. The following few
sections will try to explain the semantic of some of the L4
specific types and ideas on how to represent them using other 
methods then currently used ones ({\tt flexpage} type).

\subsection{\tt flexpage}
\label{sec:flexpage}
A flexpage is a memory region consisting of at least one 
memory frame which can be transferred from one address space
to another (or rather the rights to access it). It consists
of a base address, a size, and the transferred rights, including
the ownership. There also apply restrictions to the address ---
it has to be page size aligned --- and the size of a flexpage.
To express this using attributes, we could write
something like:
\begin{verbatim}
[in, memory, size_is(pages), 
rights(access)] void* base_address, 
[in] unsigned int pages,
[in] unsigned int access
\end{verbatim}
This scheme would add two new attributes (\verb|memory| and
\verb|rights|) and require the user to know all necessary
parameters. This can be simplified by using a predefined 
type, such as:
\begin{verbatim}
typedef struct {
[memory, size_is(pages), rights(access)]
void *base_address;
unsigned int pages;
unsigned int access;
} flexpage;
\end{verbatim}
An IDL compiler should map this constructed type 
onto the existing \verb|l4_fpage_t| type of L4 and use it
appropriately.
Since one of the goals of the IDL compiler is the usage of 
existing data types --- such as \verb|l4_fpage_t| --- this would
be contra productive.

Therefore we decided to use the special type \verb|flexpage| or
\verb|fpage|. It is directly mapped to the \verb|l4_fpage_t| type
and implies the above mentioned characteristics of a memory
region. It is simpler in the usage, because the user does
not have to transfer the L4 type to the predefined IDL type.

%TODO: make dice recognize l4_fpage_t

\subsection{Interrupts}
One of the future goals is to make the interaction with the
kernel transparent to the user as well. One could image to
describe the kernel interface using an IDL file, which defines
all type and system calls and the IDL compiler generates the
appropriate code to invoke the system calls.

This idea involves the abstraction of interrupt IPC. Therefore
an IDL type for interrupts should exist, which can be used to
describe an interrupt IPC.

{\it These are abstract ideas, which have not been implemented
and will not be in the near future.}

