\documentclass[a4paper, 10pt, twocolumn]{article}

\usepackage{epsfig}
\usepackage[latin1]{inputenc}

\usepackage{makeidx}
\usepackage{graphicx}
\usepackage{html}
\usepackage{verbatim}

%\begin{latexonly}
\usepackage{fancyvrb}
\usepackage{syntax}
\usepackage{array}
\usepackage{todo}
%\end{latexonly}

\setlength{\parindent}{0cm}
\setlength{\parskip}{0.2cm plus 0.1cm minus .1cm}

\usepackage{times}
\sloppy
\pagestyle{empty}

\begin{document}
\thispagestyle{empty}

% create a new length, which can be used as a base-length for columns
% of a table.
% arguments: #1 - the name of the length
%            #2 - number of columns of the table
%
% The length is calculated as \columnwidth-2*\tabcolsep*#2
%
\newcommand{\maketablength}[2]{%
  \newlength{#1}\setlength{#1}{-#2\tabcolsep}%
  \addtolength{#1}{#1}\addtolength{#1}{\columnwidth}%
}

% Create a table, whose first column contains text,
% and the second column fills the rest. Both columns contain parboxes.
%
% arguments: #1 - text with the maximum width of the first column
\newlength{\descriptiontablelen}
\newlength{\descriptiontabletextlen}
\newenvironment{descriptiontable}[1]{%
  %begin{latexonly}
  \settowidth{\descriptiontabletextlen}{#1}%
  \setlength{\descriptiontablelen}{-4\tabcolsep}%
  \addtolength{\descriptiontablelen}{\columnwidth}%
  \addtolength{\descriptiontablelen}{-\descriptiontabletextlen}%
  \begin{tabular}{p{\descriptiontabletextlen}p{\descriptiontablelen}}%
  %end{latexonly}  
  \html{\begin{tabular}{ll}}%
}{\end{tabular}}
  
% Create a framed table, whose first column contains text,
% and the second column fills the rest. Both columns contain parboxes.
% We can use \# to end a line. In the current version, this separates
% all rows by single lines. We cheat on latex2html by preprocessing
% the .tex file substituting the \# with a \\\hline.
%
% arguments: #1 - text with the maximum width of the first column
\newenvironment{parameterlist}[1]{%
  %begin{latexonly}
  \renewcommand{\#}{\crcr\hline}
  \settowidth{\descriptiontabletextlen}{\tt\small{}#1}%
  \setlength{\descriptiontablelen}{-4\tabcolsep}%
  \addtolength{\descriptiontablelen}{\columnwidth}%
  \addtolength{\descriptiontablelen}{-\descriptiontabletextlen}%
  \begin{tabular}{|>{\tt\small}p{\descriptiontabletextlen}|p{\descriptiontablelen}|}%
  %end{latexonly}  
  \html{\begin{tabular}{>{\tt{}}ll}}%
  \hline%
}{%
  %begin{latexonly}
   \end{tabular}%
  %end{latexonly}
  \html{\end{tabular}}}

%begin{latexonly}
% We use the '§' character to write a '$' without entering math-mode
\DeclareInputText{167}{\$}

% We use '«' and '»' to begin and end code-pieces
\renewcommand{\guillemotleft}{\begin{code}{}}
\renewcommand{\guillemotright}{\end{code}}
%end{latexonly}  
\newenvironment{code}{\tt}{}

\newcommand{\llinux}{\mbox{L\makebox[.4\width]{$^4~$}Linux}}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% The story starts here.
%
% Character-encoding, we do it the following way:
%
% (Because that stupid latex2html does not support shortverb, we do
% not use the shorthand |text| to print the text verbose. This would allow
% easy writing of variables and text while disallowing line breaking.)

% For tables, we use «text», which expands to \code{text}, which expands
% to {\tt{}text}. We cheat on latex2html by preprocessing the .tex-file.
%
% During the hole text, we can use '_' to write an underscore. We use
% '§' to write a '$' in non-verbose mode.
  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{DROPS Building Infrastructure Proposal}
\author{Jork Löser --- May 2002}
\date{}
\maketitle
\thispagestyle{empty}

\begin{abstract}
  
  To ease the writing of DROPS components, we wrote a document
  defining the unified infrastructure of the DROPS source tree. We
  implemented a set of macros and some tools providing the described
  functionality.

\end{abstract}

\section{History}

In 2000, some people (see Section \ref{sec:acknowledgements})
began to write down their expectations to a build system and defined its
behavior. A macro-system implementing this behavior was implemented
and is ready to use since then \cite{cvs:dropsmk}. However, there were
some features missing, such as support for an easy IDL compilation.
Since then, the macros were modified and adapted a little bit, but
they are still unused.

Within the last two years, the structure of the building system
changed and today, we have a quite clean source tree with a lot of
unwritten rules. Hence, writing a Makefile still bears some
uncertainties. Users fall back to copy-and-paste usually, accompanied
by praying or just hoping depending on personal taste.

We think, the macro-system written two years ago is not used for two
reasons: Firstly, it was not well documented. There is an ASCII file
describing some of the magic, but there is not everything in it.
Secondly, the availability of the macro-system was not communicated.
Especially for students, it should have been a big help: Writing 5
lines and getting every functionality we ever thought about should be
much easier than reinventing all that stuff over and over again.

To sum it up, we give it another try. This time, there is a spec
\cite{www:buildingInfra}, 29~pages currently. It is comprising as
there is nearly every functionality in it our DROPS system currently
offers (and a bit more). We communicate that there is help for writing
Makefiles. Finally, migrating to the proposed system can be done on a
directory-by-directory base, placing no burden for keeping old code.

\section{Features}
\subsection{Roles}
\label{sec:rolefiles}

Each directory in the source-tree has a distinct task, i.e., each
directory builds special kinds of targets. This disctinct task is
expressed by a role, while each role is one of the following:

\begin{descriptiontable}{include:}
subdir:  & be a container for other directories\\
prog:    & build executable binaries \\
lib:     & build libraries \\
idl:     & translate IDL definitions into code \\
include: & hold include-files which have to be installed \\
\end{descriptiontable}

To define the role of a directory, the Makefile of that directory has
to include an according Makefile-include, the \emph{role-file}. The
role-files are templates that define rules and variables to build the
proper targets. Make-variables are used to control the behaviour of
the role-files. Each Makefile must include exactly one role-file.

As usual, defining the Make-variables to control the role-files can
happen in 4 different places: The Makefile, a local «Makeconf.local»,
a packet-local «Makeconf.local», and a global «Makeconf.local».

\subsection{Provided Make-Targets}

With all roles, the following phony targets can be built using «make
<target>».

\begin{description}
%\setlength{\parskip}{0pt}
\item [all::]           The default target. Builds binaries,
                        libraries, generating .c and .h files from
                        IDLs. It depends on the role, what exactly is done.
\item [\{txt, old,\}config::]
                        Run the DROPS Configuration Tool. See
                        Section \ref{sec:config} for details on this.
\item [install::]       Installs the generated files into the
                        installation-tree «§(DROPS_STDDIR)».
\item [clean::]         Deletes all intermediate files generated
                        during compilation. This does \emph{not}
                        delete the generated targets such as binaries
                        or libraries.
\item [cleanall::]      Deletes all generated files. Depends on
                        \textbf{clean::}.
\item [help::]          Displays a short overview of the make-targets
                        available with the role.
\end{description}
                
\subsection{DROPS Configuration Tool}
\label{sec:config}

Most role-files provide support for an interactive configuration tool,
which is derived from the \emph{config} and \emph{menuconfig} tools of
Linux. It differs in that it can be configured using environment
variables. Some minor extensions are build in too.

The configuration tool is run if the user runs ``«make config»'',
``«make txtconfig»'' or ``«make oldconfig»''.

As within Linux, the configuration tool creates two files: One to be
included within the Makefile itself. This file also stores the
configuration. The other generated file is an include file to be
included in C source files. It is ensured that a default configuration
file and a default configuration include file is created prior to
compiling other source files. Do we have to mention that you can
define every filename according to your personal taste, even
«__many_underscores__in_name___»?\footnote{Though we do not recommend
  this example}


\subsection{Target-specific variables}

With all building roles, there can be multiple targets. This is, you
can build multiple binaries or libraries within one directory. This
requires configuration variables to be interpreted on a per-target
base. For an example, we look at the «SRC_C» variable. With the prog
role, this variable defines which .c-files go into a binary. If you
want the two binaries {\bf{}A} and {\bf{}B} to be created in the
directory, with {\bf{}A} having «a.c» and «x.c» compiled in, and
{\bf{}B} having «b.c» and «x.c» compiled in, specify «SRC_C=x.c»,
«SRC_C_A=a.c» and «SRC_C_B=b.c». Further specify «TARGET=A B», and
that's it.

\subsection{Dependencies}

Dependencies are built automatically. Each source-file is accompanied
by an according «.x.d» file with x being replaced by the original
filename. Typically these files contain header files. Also, each
binary is accompanied by an according «.y.d» file with y being
replaced by the binary filename. These files typically contain
libraries. Dependency files are created automatically and they are
used automatically. Removing a dependency (e.g., an include-file),
does not harm the build process. If this is not the case, the
build-system has a bug.


\subsection{Relocation of Binaries}
\label{sec:reloc}

We support two flavors of finally linking: Relocating a formerly
relocatable binary or build relocated binaries in a normal link
process. With the former method, all binaries a built relocatable
first. These relocatable binaries have all symbols resolved and all
required libraries linked, but the final start-address is not yet
defined. In a second step, invoked with «make reloc», a program
determines the memory ranges required by each binary and links them to
different but contiguous locations.

The other method is the approach using the STATIC file. A default link
address must be specified in the Makefile.

\subsection{IDL files}

The purpose of the IDL role is to translate IDL definition files into
appropriate .c- and .h-files. Different compilers are supported and
the user needs not to take care of the resulting filenames. To compile
the generated files, directories with the prog or lib role are used.
These import the generated files using IDL-related parameters.

\subsection{Profiling and Shared libraries}

Compiling profiling versions of libraries and binaries is easy: The
user just has to specify one target-specific flag: BUILD_PROFILE. As a
result, profiling versions of objects, libraries and binaries are
built additionally to the normal ones.

Building shared libraries is easy too, the build-parameter is
BUILD_SHARED resp. BUILD_PIC.

\subsection{Installation}

When includes, libraries or binaries are built, they are installed
into a \emph{local installation directory}. This local installation
directory is a parameter of the according role-file and is
«§(L4DIR)/include» for includes, «§(L4DIR)/bin» for binaries and
«§(L4DIR)/lib» for libraries per default. The commands used for
installation can be selected as well. Includes and libraries use
symbolic links, binaries are copied and stripped.

Using «make install» installs includes, libraries and binaries into
the \emph{global installation directory}, aka «/home/drops». Of
course, no symbolic links are used here.


\section{Open Issues}

\subsection{Architectures and Modes}

Currently, there is something like a support for different
architectures: We allow different sets of source-files and options
depending on the architecture to build. When building for multiple
architectures from one source-tree, each architecture is built into a
separate subdirectory. However, for compatibility reasons all binaries
and libraries are installed into the same installation directories.
Hence, this support is really suboptimal.

There is no final consensus on how to support different target
systems. This is, the same hardware platform and the same kernel, but
different execution environments. We see the need to support the
following ones:

\begin{description}

\item[oskit10]

  aka «oskit10_l4env_full». This mode allows to use drivers and the
  infrastructure from the Flux OSKit 0.97, together with the L4Env
  \cite{www:l4env}. Binaries are linked against the huge freebsd-libc.

\item[loader]

  Like oskit10 mode, but libraries are dynamically linked. Binaries
  must be loaded with the L4Env loader \cite{www:l4envloader}.

\item[l4env]

  aka «oskit10_l4env_tiny». This mode allows to use some basic libc
  functionality from the Flux OSKit 0.97, together with the L4Env.
  Binaries are linked against the small mini-libc «libmc».

\item[sigma0]

  This has no support for L4Env, the result is a plain L4 application.
  Binaries are linked against the small mini-libc «libmc» from the
  Flux OSKit 0.6.

\item[l4linux]

  Use this mode to build \llinux{} user-mode applications with L4Env
  support. Binaries are linked statically against special L4Env
  libraries and dynamically against standard-libc.


\item[l4linux-kern]

  Use this mode to build \llinux{} kernel-mode modules with L4Env
  support. Modules are linked against special L4Env libraries.

\item[host]

  Build an application for the host operating system.

\end{description}

\subsection{"doc" Role}

A role for documentation might be handy. This should have a nice
environment for doxygenned documentation and should provide support
for latex, xdvi, postscript and PDF generation. We also have to unify
how to publish this documentation on the web.

\begin{description}
\setlength{\parskip}{0pt}

\item[Request from Lars:] Allow to define the target directory in the
  global Makeconf.local.

\item[Answer:] Define a variable, e.g., «§(HTML_DOC_DIR)» specifying
were to install the html-files. This could then be set to
«§(HOME)/public_html/docs/§(PKGNAME)/».

\end{description}


\section{Acknowledgements}
\label{sec:acknowledgements}

This document is based on the work of Lukas Grützmacher, Michel
Hohmuth, Jork Löser and Lars Reuther done in 2000. Since then, Frank Mehnert
proposed a lot of functionality which is included in this document as well.

{ %begin{latexonly}
    \section{\refname}
    \renewcommand{\section}[2]{}
    \footnotesize
  %end{latexonly}
  \bibliographystyle{plain} %alpha}
  \bibliography{own}
}
\section{Examples}

%\newcommand{\exmpl}{1}{\begin{minipage}{\columnwidth}%
%      \VerbatimInput{#1}\end{minipage}}

Here we show some Makefile examples. As a base, we used the names
package and modified its Makefiles. The names package has the
following subtree structure:

\newcommand{\vfontsize}{\small}
\newcommand{\vframe}{single}

\begin{minipage}{\columnwidth}
\VerbatimInput[xleftmargin=3ex]{code/names-tree}
\end{minipage}

The Makefiles for the subdirectories names/, examples/, lib/ and
server/ only differ in the values for «§PKGDIR», we only list one.

\noindent\begin{minipage}{\columnwidth}
\VerbatimInput[fontsize=\vfontsize,
               frame=\vframe, label=Subdirs]
               {code/names-subdir}
\end{minipage}

\noindent\begin{minipage}{\columnwidth}
\VerbatimInput[fontsize=\vfontsize,
               frame=\vframe, label=examples/demo/Makefile]
               {code/names-examples-demo}
\end{minipage}

\noindent\begin{minipage}{\columnwidth}
\VerbatimInput[fontsize=\vfontsize,
               frame=\vframe, label=examples/lister/Makefile]
               {code/names-examples-lister}
\end{minipage}

\noindent\begin{minipage}{\columnwidth}
\VerbatimInput[fontsize=\vfontsize,
               frame=\vframe, label=lib/src/Makefile]
               {code/names-lib-src}
\end{minipage}

\noindent\begin{minipage}{\columnwidth}
\VerbatimInput[fontsize=\vfontsize,
               frame=\vframe, label=server/src/Makefile]
               {code/names-server-src}
\end{minipage}

The server/include subdirectory contains no Makefile. We did not look
at the Makefile in the man/ subdirectory.

For comparison, the original Makefile from the server/src/ directory
is printed here:

\begin{minipage}{\columnwidth}
\VerbatimInput[fontsize=\tiny,
               frame=\vframe, label=Original server/src/Makefile]
               {code/names-orig-server-src}
\end{minipage}

% align the last column
\pagebreak

\end{document}
% Localwords: CVS unset
% Local variables:
%  compile-command: "make"
% End:
